---
title: "risksetROC_practice"
output: html_document
---

```{r library}
library(risksetROC)
```

# R package 'risksetROC'

## Mayo PBC data
418 obs. of 20 variables  
*age*:	in years  
*albumin*:	serum albumin (g/dl)  
*alk.phos*:	alkaline phosphotase (U/liter)  
*ascites*:	presence of ascites  
*ast*:	aspartate aminotransferase, once called SGOT (U/ml)  
*bili*:	serum bilirunbin (mg/dl)  
*chol*:	serum cholesterol (mg/dl)  
*copper*:	urine copper (ug/day)  
*edema*:	0 no edema, 0.5 untreated or successfully treated 1 edema despite diuretic therapy  
*hepato*:	presence of hepatomegaly or enlarged liver  
*id*:	case number  
*platelet*:	platelet count  
*protime*:	standardised blood clotting time  
*sex*:	m/f  
*spiders*:	blood vessel malformations in the skin  
*stage*:	histologic stage of disease (needs biopsy)  
*status*:	status at endpoint, 0/1/2 for censored, transplant, dead  
*time*:	number of days between registration and the earlier of death, transplantion, or study analysis in July, 1986  
*trt*:	1/2/NA for D-penicillmain, placebo, not randomised  
*trig*:	triglycerides (mg/dl)  
```{r}
# data load
data(pbc)
str(pbc)
head(pbc)
```
