---
title: "risksetROC_practice"
output: html_document
---

```{r library}
library(risksetROC)
```

# R package 'risksetROC'

## Mayo PBC data
A data frame with 418 obs. of 20 variables      
This is Mayo PBC data as obtained from the website: http://lib.stat.cmu.edu/datasets/pbc   

*id*: patient id 
*fudays*: follow-up days, number of days between registration and the earlier of death, transplantation, or study analysis time in July, 1986   
*status*: survival status  
*drug*: 1 = D-penicillamine, 2 = placebo  
*age*: age in days  
*sex*: 0 = male, 1 = female   
*ascites*: presence of asictes: 0=no 1=yes  
*hepatom*: presence of hepatomegaly: 0=no 1=yes   
*spiders*: presence of spiders: 0=no 1=yes   
*edema*: presence of edema: 0=no edema and no diuretic therapy for edema; .5 = edema present without diuretics, or edema resolved by diuretics; 1 = edema despite diuretic therapy   
*bili*: serum bilirubin in mg/dl   
*chol*: serum cholesterol in mg/dl  
*albumin*: albumin in gm/dl  
*copper*: urine copper in ug/day    
*alkphos*: alkaline phosphatase in U/liter  
*sgot*: SGOT in U/ml    
*trig*: triglicerides in mg/dl   
*platelet*: platelets per cubic ml / 1000   
*protime*: prothrombin time in seconds   
*stage*: histologic stage of disease    
```{r}
# data load
data(pbc)
str(pbc)
head(pbc)
```

for(i in 1:length(pbc)){
  hist(pbc[,i], main=paste("pbc","[",i,"]"))

범주형 자료는 as.factor로 변환한다. 
```{r}
## id, status, drug, sex, ascites, hepatom, spiders, edema, stage(순서 있음)
#pbc$id <- as.factor(pbc$id)
#pbc$status <- as.factor(pbc$status)
#pbc$drug <- as.factor(pbc$drug)
#pbc$sex <- as.factor(pbc$sex)
#pbc$hepatom <- as.factor(pbc$hepatom)
#pbc$spiders <- as.factor(pbc$spiders)
#pbc$edema <- as.factor(pbc$edema)
```


```{r}
# histogram
# scale 변환해야 할 변수를 알아본다. 
colnames(pbc)
 #[1] "id"       "fudays"   "status"   "drug"     "age"      "sex"      "ascites"  "hepatom"  "spiders"  "edema"
 #[11] "bili"     "chol"     "albumin"  "copper"   "alkphos"  "sgot"     "trig"     "platelet" "protime"  "stage"

for(i in 1:ncol(pbc)){
  hist(pbc[,i], xlab='', main=paste("pbc","[",i,"]"))
}
```
11,14,15번째 변수(각 bili, copper, alkphos)는 스케일이 치우쳐 있으므로 log 변환을 하면 좋다.   
논문에서는 bili, protime, edema, albumin, age 변수를 marker의 base covariates로 사용하였다.  
논문의 저자는 bili뿐만 아니라 19번째 변수 protime또한 log 변환을 해주어 사용하였다. 
   
   
pbc data 속 앞 312명의 patients는 randomized patient다. 따라서 앞 312명의 환자만을 자료로 사용하여 분석한다.  
```{r}
unique(pbc$id)   #id: 1~418
```
    
```{r}
pbc1 <- pbc[1:312,]    #pbc1: id1~312 환자만 불러온 데이터 
table(pbc1$status)   #0:censored , 1:transplant , 2:dead   
```
pbc 데이터는 '다상태' 자료이다. 이번 분석에서는 다상태가 아닌 2가지 상태만으로 변경하여 사용한다.  
즉, censored와 transplant를 모두 'censored'로 간주하고, dead는 그대로 'dead(event occur)'로 간주한다.   
```{r}
survival.status <- ifelse(pbc1$status==2,1,0)    #survival.status: status
table(survival.status)
```
```{r}
survival.time <- pbc1$fudays    #survival.time: follow-up days
pbc1$status1 <- survival.status    #add 'survival.status' in pbc1 dataset 
```
  
  
### model fitting
```{r}
fit <- coxph(Surv(fudays,status1)~log(bili)+log(protime)+edema+albumin+age, data=pbc1)
eta <- fit$linear.predictors
```




