---
title: "risksetROC_practice"
output: html_document
---

```{r library}
library(risksetROC)
```

# R package 'risksetROC'

## Mayo PBC data
A data frame with 418 obs. of 20 variables      
This is Mayo PBC data as obtained from the website: http://lib.stat.cmu.edu/datasets/pbc   

*id*: patient id 
*fudays*: follow-up days, number of days between registration and the earlier of death, transplantation, or study analysis time in July, 1986   
*status*: survival status  
*drug*: 1 = D-penicillamine, 2 = placebo  
*age*: age in days  
*sex*: 0 = male, 1 = female   
*ascites*: presence of asictes: 0=no 1=yes  
*hepatom*: presence of hepatomegaly: 0=no 1=yes   
*spiders*: presence of spiders: 0=no 1=yes   
*edema*: presence of edema: 0=no edema and no diuretic therapy for edema; .5 = edema present without diuretics, or edema resolved by diuretics; 1 = edema despite diuretic therapy   
*bili*: serum bilirubin in mg/dl   
*chol*: serum cholesterol in mg/dl  
*albumin*: albumin in gm/dl  
*copper*: urine copper in ug/day    
*alkphos*: alkaline phosphatase in U/liter  
*sgot*: SGOT in U/ml    
*trig*: triglicerides in mg/dl   
*platelet*: platelets per cubic ml / 1000   
*protime*: prothrombin time in seconds   
*stage*: histologic stage of disease    
```{r}
# data load
data(pbc)
str(pbc)
head(pbc)
```

범주형 자료는 as.factor로 변환한다. 
```{r}
# id, status, drug, sex, ascites, hepatom, spiders, edema, stage(순서 있음)
pbc$id <- as.factor(pbc$id)
pbc$status <- as.factor(pbc$status)
pbc$drug <- as.factor(pbc$drug)
pbc$sex <- as.factor(pbc$sex)
pbc$hepatom <- as.factor(pbc$hepatom)
pbc$spiders <- as.factor(pbc$spiders)
pbc$edema <- as.factor(pbc$edema)
```


```{r}
# histogram
# scale 변환해야 할 변수를 알아본다. 
colnames(pbc)

for(i in 1:length(pbc)){
  hist(pbc[,i], main=paste("pbc","[",i,"]"))
}
```

